### 2015-06-13 writing/S'15 LSTM/abstract.tex
Paper push.

\begin{abstract}
Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but these models largely have not outperformed simpler sequence-based models by substantial margins. We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured are built around, mitigating the advantage of the tree-structured models. We investigate this possibility by evaluating both models on an artificial talk for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though that it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
\end{abstract}

>>>>

\begin{abstract}
Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but these models largely have not outperformed simpler sequence-based models by substantial margins. We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured are built around, mitigating the advantage of the tree-structured models. We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though that it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
\end{abstract}

### 2015-06-13 writing/S'15 LSTM/discussion.tex
Cleanup, citations, etc.
Paper push.

\begin{figure*}[t!]
  \centering
\begin{subfigure}[t]{0.29\textwidth}
  \includegraphics[height=1.6in]{fig3.pdf}
  \caption{Training expression size $\le$3.}
  \end{subfigure}~~
\begin{subfigure}[t]{0.29\textwidth}
    \includegraphics[height=1.6in]{fig4.pdf}
  \caption{Training expression size $\le$4.}
  \end{subfigure}~~
\begin{subfigure}[t]{0.4\textwidth}
      \includegraphics[height=1.6in]{fig6.pdf}
  \caption{Training expression size $\le$6.}
\end{subfigure}
  \caption{Results on three experiments with increasingly rich training sets. The horizontal axis on each graph divides the test set data points (expression pairs) into bins by the number of logical operators in the more complex of the two expressions in the pair.}
  \label{prop-results} 
\end{figure*}

The results are shown in Figure \ref{prop-results}. All three models were able to largely memorize the training data (TODO, TODO, and TODO for $\le$6). With sufficient training data (in the $\le$6 setting), the two tree models were able to perform at above 99\% (TODO) on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets.

@@

We find that

>>>>

The results are shown in Figure \ref{prop-results}. All three models were able to largely memorize the training data (TODO, TODO, and TODO for $\le$6). With sufficient training data (in the $\le$6 setting), the two tree models were able to perform at above 99\% (TODO) on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets. These robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.

@@

We find that all three \todo{Still using three?} models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree structured model's biases allow them to learn to do this more effectively from less data. 
We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension. We finally suggest that there is value in attempting to further develop this ability in sequence models without fully sacrificing the flexibility that makes these models succeed.

### 2015-06-13 writing/S'15 LSTM/figure1.tex
Paper push.

\begin{subfigure}[t]{0.45\textwidth}
\centering
\scalebox{0.75}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{29pt}

    \tikzstyle{label}=[text width=40mm,align=center]    
    \tikzstyle{softmax}=[fill=red!40,text width=40mm,align=center]
    \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center]
    \tikzstyle{e}=[fill=green!40,text width=26mm,align=center]
    \tikzstyle{m}=[draw=black,text width=38mm,align=center]

@@

\begin{subfigure}[t]{0.45\textwidth}
  \centering
\scalebox{0.75}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{29pt}

@@

     \caption{The architecture for the TreeRNN and TreeRNTN sentence models. Terminal nodes are learned embeddings and nonterminal nodes are either NN or NTN layers with $\tanh$ nonlinearities.}\label{fig:model:tree}

@@

\begin{subfigure}[t]{0.45\textwidth}
  \centering
\scalebox{0.75}{
 \begin{tikzpicture}
    \def\dx{35pt}
    \def\dy{29pt}

>>>>

\begin{subfigure}[t]{0.45\textwidth}
\centering
\scalebox{0.75}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{27pt}

    \tikzstyle{label}=[text width=40mm,align=center,text height=2mm]    
    \tikzstyle{softmax}=[fill=red!40,text width=40mm,align=center,text height=2mm]
    \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center,text height=2mm]
    \tikzstyle{e}=[fill=green!40,text width=26mm,align=center,text height=2mm]
    \tikzstyle{m}=[draw=black,text width=38mm,align=center,text height=2mm]

@@

\begin{subfigure}[t]{0.45\textwidth}
  \centering
\scalebox{0.75}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{27pt}

@@

     \caption{The architecture for the tree-structured sentence models. Terminal nodes are learned embeddings and nonterminal nodes are NN or NTN layers.}\label{fig:model:tree}

@@

\begin{subfigure}[t]{0.45\textwidth}
  \centering
\scalebox{0.75}{
 \begin{tikzpicture}
    \def\dx{35pt}
    \def\dy{27pt}

### 2015-06-13 writing/S'15 LSTM/methods.tex
Paper push.

\begin{figure}[t!]
  \centering
  \input{figure1}
  \caption{In our model, two separate sentence models---based on either tree (b) or sequence (c) models---build up vector representations for each of two sentences. A multilayer classifier component (a) then uses the resulting vectors to predict a label that reflects the logical relationship between the two sentences.} 
  \label{sample-figure}
\end{figure}

\begin{figure*}[t]
  \centering
\begin{subfigure}[t]{0.29\textwidth}
  \includegraphics[height=1.6in]{fig3.pdf}
  \caption{Training expression size $\le$3.}
  \end{subfigure}~~
\begin{subfigure}[t]{0.29\textwidth}
    \includegraphics[height=1.6in]{fig4.pdf}
  \caption{Training expression size $\le$4.}
  \end{subfigure}~~
\begin{subfigure}[t]{0.4\textwidth}
      \includegraphics[height=1.6in]{fig6.pdf}
  \caption{Training expression size $\le$6.}
\end{subfigure}
  \caption{Results on three experiments with increasingly rich training sets. The horizontal axis on each graph divides the test set data points (expression pairs) into bins by the number of logical operators in the more complex of the two expressions in the pair.}
  \label{prop-results} 
\end{figure*}

We use the model architecture depicted in Figure \ref{fig:model:top}. This approach is loosely based on the model used in the experiments by \newcite{Bowman:Potts:Manning:2014}. The model architecture uses two copies of a single sentence embedding network---a TreeRNN, TreeRNTN, or LSTM RNN, depending on the precise experiment---to embed the premise and hypothesis sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations. Since the embeddings are computed separately, the embedding models are forced encode every element of the sentence meaning that the downstream model will need. Though this prevents us from exploiting cross-sentence alignment information that could allow for the learning of more effective entailment models, it allows us to focus our evaluation on models that capture the meanings of single sentences.

\paragraph{Classifier}
The classifier component of the model consists of a combinging layer which takes the two sentence representations as inputs, followed by two neural network layers, then a softmax classifier.
For the combining layer, we use a (recursive) neural tensor network ((R)NTN, \citealt{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:
\begin{gather} 
\label{TreeRNN}
\vec{y}_{\textit{RNN}} = \tanh(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
\label{TreeRNTN} 
\vec{y}_{\textit{RNTN}} = \vec{y}_{\textit{RNN}} + \tanh(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather} 
Our model is largely identical to the Bowman et al. model, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence embedding models are compared in as similar a setting as possible.

\paragraph{Sentence embedding}
The sentence embedding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector. We experiment with two tree-structured models (Figure \ref{fig:model:tree}), one with an RNTN activation function or a plain RNN activation function. Both are structured using the parses that are provided with the data. In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} with the precise formulation from \newcite{zaremba2015recurrent}.

We additionally experimented with a plain RNN---the sequence equivalent of the TreeRNN---but were not able to coerce that weaker model into learning to model even the training data well, so they are omitted below. LSTM, in contrast are the most widely used of a class of models that attempt to avoid the extreme sensitivity to initialization and gradient magnitude that make RNN training so unreliable.

\paragraph{Training} We randomly initialize all embeddings and layer parameters, and then train all parameters using minibatch stochastic gradient descent with AdaDelta \cite{zeiler2012adadelta} learning rates. Our objective is the standard negative log likelihood classification objective with L2 regularization. All models were stopped at 100 epochs (TODO), it which point all had largely converged without significantly declining from their peak performances.

>>>>

\begin{figure}[t]
  \centering
  \input{figure1}
  \caption{In our model, two copies of a sentence model---based on either tree (b) or sequence (c) models---encode the two sentences. A multilayer classifier component (a) then uses the resulting vectors to predict a label that reflects the logical relationship between the two sentences.} 
  \label{sample-figure}
\end{figure}

\begin{figure*}[t]
  \centering
  \begin{subfigure}[t]{0.04\textwidth}
      \includegraphics[height=1.45in]{scale.pdf}
\end{subfigure}
\begin{subfigure}[t]{0.24\textwidth}
  \includegraphics[height=1.45in]{fig3c.pdf}
  \caption{Training on sz. $\le$3.}
  \end{subfigure}~~~
\begin{subfigure}[t]{0.24\textwidth}
    \includegraphics[height=1.45in]{fig4c.pdf}
  \caption{Training on sz. $\le$4.}
  \end{subfigure}~~~~
\begin{subfigure}[t]{0.24\textwidth}
      \includegraphics[height=1.45in]{fig6c.pdf}
  \caption{Training on sz. $\le$6.}
\end{subfigure}~~
\begin{subfigure}[t]{0.08\textwidth}
      \includegraphics[height=1.45in]{leg.pdf}
\end{subfigure}
  \caption{Results on three experiments with increasingly rich training sets. The horizontal axis on each graph divides the test set data points (expression pairs) into bins by the number of logical operators in the more complex of the two expressions in the pair.}
  \label{prop-results} 
\end{figure*}

We use the model architecture depicted in Figure \ref{fig:model:top}. This approach builds on the model used in \newcite{Bowman:Potts:Manning:2014}. The model architecture uses two copies of a single sentence embedding network---a TreeRNN, TreeRNTN, or LSTM RNN, depending on the precise experiment---to embed the premise and hypothesis sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations. Since the embeddings are computed separately, the embedding models are forced to encode every element of the sentence meaning that the downstream model will need.

\paragraph{Classifier}
The classifier component of the model consists of a combinging layer which takes the two sentence representations as inputs, followed by two neural network layers, then a softmax classifier.
For the combining layer, we use a (recursive) neural tensor network ((R)NTN, \citealt{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:
\begin{gather} 
\label{TreeRNN}
\vec{y}_{\textit{RNN}} = \tanh(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
\label{TreeRNTN} 
\vec{y}_{\textit{RNTN}} = \vec{y}_{\textit{RNN}} + \tanh(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather}

Our model is largely identical to the Bowman et al.~model, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence embedding models are compared in as similar a setting as possible.

\paragraph{Sentence embedding}
The sentence embedding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector. We experiment with two tree-structured models (Figure \ref{fig:model:tree}), one with an RNTN activation function or a plain RNN activation function. Both are structured using the parses that are provided with the data. In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}.\todo{Okay to omit negative result on sequence RNN?}

\paragraph{Training} We randomly initialize all embeddings and layer parameters, and then train all parameters using minibatch stochastic gradient descent with AdaDelta \cite{zeiler2012adadelta} learning rates. Our objective is the standard negative log likelihood classification objective with L2 regularization. All models were stopped at 100 epochs (TODO), at which point all had largely converged without significantly declining from their peak performances.
